# Data
This project uses only one dataset that was collected 'https://www.climatewatchdata.org/' by Krishen Jagani and Rui Wang.

## Description
The dataset was imported in a csv format. The dataset tracks emission of greenhouse gases across different sectors and gases over time, allowing for a detailed analysis
of historical trends. The dataset has the following categorical variales and contionous variables:

**Categorical Variables**: Country (`ISO`, `Country`), data source, sector, gas type, and unit.

**Continuous Variables**: Yearly emissions values (1990–2021).

**Dimensions**: 8,214 rows x 37 columns.

**Frequency of Updates**: Likely annual.

**Issues**: Missing values for earlier years and specific sectors. Moroever, the data is aggregated at varying levels (like global, country, sector, gases)
which may filtering for specific analysis.

## Missing value analysis


```{r}
emissions_data <- read.csv("C:\\Users\\HP WORLD\\Downloads\\historical_emissions (1)\\historical_emissions.csv",na="N/A")
#head(emissions_data)
```

```{r,fig.width=11,fig.height=8}
library(ggplot2)
library(dplyr)
library(tidyverse)

#  missing values per column
missing_value_counts <- colSums(is.na(emissions_data))
missing_df <- data.frame(Variable = names(missing_value_counts), Missing = missing_value_counts)

# Bar plot 
ggplot(missing_df, aes(x = reorder(Variable, Missing), y = Missing)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Count of Missing Values by Variable",
    x = "Year",
    y = "Number of Missing Values"
  ) +
  theme_minimal(base_size = 16) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
# Key observations from this plot:
1. Variables representing earlier years have higher number of missing values. This could indicate data collection during those times were less consistent
or non existent. However, the proportion of missing values is very less. As we can see that year 1990 has the highest amount of missing values(around 250). However, the 
dataset has 8214 rows, which makes the proportion of missing values small. 
2. Another trend that we observe, is the decrease in number of missing values with time as recent years have less number of missing values. This might suggest improvements
in data reporting and collection over the years.
3. The key variables - Country, Gas, Sector and Unit have no missing values.

```{r,fig.width=15}
# Calculate missing values for each gas 
missing_by_gas <- emissions_data |>
  pivot_longer(cols = starts_with("X"), names_to = "Year", values_to = "Value") |>
  group_by(Gas, Year) |>
  summarise(Missing_Count = sum(is.na(Value)), .groups = "drop")

#  Facets for gases
ggplot(missing_by_gas, aes(x = Year, y = Missing_Count)) +
  geom_col(fill = "blue") +
  facet_wrap(~ Gas) +
  labs(
    title = "Missing Values Faceted by Gas",
    x = "Year",
    y = "Number of Missing Values"
  ) +
  theme_minimal(base_size = 16) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
# Key Observations from this plot
1. Carbon Dioxide (CO2) has the most amount of missing values in year columns. Methane (CH4) and Nitrous Oxide (N2O) have
relatively same number of missing values over the years. Whereas, F-Gas has no missing values.
