# Results

```{r}
library(readr)
library(tidyverse)
```

```{r}
data <- read.csv("Datasets/historical_emissions.csv", na.strings = c("NA", "N/A"))


dim(data)
#head(data)
```
We begin our analysis by cleaning the data first. This includes removing all the rows with missing values and dropping Data.source column.
Further, we also prepare a long format of the data well before, as this would be used in few graphs coming forward.
```{r}
cleaned_data <- na.omit(data)
cleaned_data <- cleaned_data[,!names(cleaned_data) %in% c("Data.source")]
dim(cleaned_data)
#str(cleaned_data)
data_long <- cleaned_data |>
  pivot_longer(
    cols = starts_with("X"),    
    names_to = "Year",          
    values_to = "Emissions"     
  ) |>
  mutate(Year = as.numeric(sub("X", "", Year)))
dim(data_long)
```
1.We first begin by creating a line plot to observe the emission trends of different gases over the years.
```{r,fig.width=15}
global_emissions_gases <- subset(data_long, Country == "World" & Sector == "Total including LUCF")
ggplot(global_emissions_gases, aes(x = Year, y = Emissions, color = Gas)) +
  geom_line(size = 1) +
  geom_point(size = 2)+
  scale_x_continuous(breaks = seq(1990, 2021, by = 1)) +
  labs(title = "Global Emissions Over Time by Gas",
       x = "Year",
       y = "Emissions (MtCO₂e)",
       color = "Gas") +
  theme(legend.position = "bottom")
```
As we see observe that only CO2 is showing a significant trend, we will apply a log scale to see if there exist any significant trends in the emission of other gases

```{r,fig.width=15}

ggplot(global_emissions_gases, aes(x = Year, y = Emissions, color = Gas)) +
  geom_line(size = 1) +
  geom_point(size = 2)+
  scale_x_continuous(breaks = seq(1990, 2021, by = 1)) +
  scale_y_continuous(trans = 'log10', labels = scales::comma) +
  labs(title = "Global Emissions Over Time by Gas",
       x = "Year",
       y = "Emissions (MtCO₂e)(on log scale)",
       color = "Gas") +
  theme(legend.position = "bottom")


```
After applying the log scale, we discover another piece of information. Along with Co2, F-gas emissions have also increased significantly recently compared to the 
earlier years.



2. Next, we attempt to get an insight into the countries that are responsible for most amount of emissions. We filter our selection by top 10 countries.
```{r}
data_2021 <- subset(data_long, Year == 2021 & Gas == "CO2" & Sector == "Total including LUCF")
top_10 <- data_2021[order(-data_2021$Emissions), ][2:11, ]

ggplot(top_10, aes(x = reorder(Country, Emissions), y = Emissions)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  scale_y_continuous(breaks = seq(0, max(top_10$Emissions), by = 1000), labels = scales::comma) +
  labs(title = "Top 10 CO2 Emitting Countries in 2021",
       x = "Country",
       y = "Emissions (MtCO₂e)") +
  coord_flip()
```
```{r}
data_2021 <- subset(data_long, Year == 2021 & Gas == "CH4" & Sector == "Total including LUCF")
top_10 <- data_2021[order(-data_2021$Emissions), ][2:11, ]

ggplot(top_10, aes(x = reorder(Country, Emissions), y = Emissions)) +
  geom_bar(stat = "identity", fill = "pink") +
  scale_y_continuous(breaks = seq(0, max(top_10$Emissions), by = 1000), labels = scales::comma) +
  labs(title = "Top 10 CH4 Emitting Countries in 2021",
       x = "Country",
       y = "Emissions (MtCO₂e)") +
  coord_flip()
```

```{r}
data_2021 <- subset(data_long, Year == 2021 & Gas == "F-Gas" & Sector == "Total including LUCF")
top_10 <- data_2021[order(-data_2021$Emissions), ][2:11, ]

ggplot(top_10, aes(x = reorder(Country, Emissions), y = Emissions)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  scale_y_continuous(breaks = seq(0, max(top_10$Emissions), by = 1000), labels = scales::comma) +
  labs(title = "Top 10 F-Gas Emitting Countries in 2021",
       x = "Country",
       y = "Emissions (MtCO₂e)") +
  coord_flip()
```

```{r}
data_2021 <- subset(data_long, Year == 2021 & Gas == "N2O" & Sector == "Total including LUCF")
top_10 <- data_2021[order(-data_2021$Emissions), ][2:11, ]

ggplot(top_10, aes(x = reorder(Country, Emissions), y = Emissions)) +
  geom_bar(stat = "identity", fill = "purple") +
  scale_y_continuous(breaks = seq(0, max(top_10$Emissions), by = 1000), labels = scales::comma) +
  labs(title = "Top 10 N2O Emitting Countries in 2021",
       x = "Country",
       y = "Emissions (MtCO₂e)") +
  coord_flip()
```
China and United States dominate the list of top emitters with India and Russia also making significant contributions.

3.Next we deep diver into the sector wise contributions of emissions among the top 10 countries emitting CO2
```{r,fig.width=15}
# Calculate LUCF emissions
top_countries <- top_10$Country
sector_data <- subset(data_long, Year == 2021 & Country %in% top_countries & Gas == "CO2")

# Create a new LUCF column by calculating the difference
sector_data_lucf <- sector_data |>
  filter(Sector %in% c("Total including LUCF", "Total excluding LUCF")) |>
  spread(Sector, Emissions) |>
  mutate(LUCF = `Total including LUCF` - `Total excluding LUCF`) |>
  gather(key = "Sector", value = "Emissions", c("LUCF", "Total excluding LUCF"))

# Combine with other sectors
sector_data_combined <- sector_data |>
  filter(!Sector %in% c("Total including LUCF", "Total excluding LUCF")) |>
  bind_rows(sector_data_lucf)

# Plot the updated data
ggplot(sector_data_combined, aes(x = Country, y = Emissions, fill = Sector)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c(
    "Building" = "#E69F00",
    "Bunker Fuels" = "#56B4E9",
    "Electricity/Heat" = "#009E73",
    "Energy" = "#F0E442",
    "Fugitive Emissions" = "#0072B2",
    "Industrial Processes" = "#D55E00",
    "Land-Use Change and Forestry" = "#CC79A7",
    "Manufacturing/Construction" = "#999999",
    "Other Fuel Combustion" = "#B2DF8A",
    "Total excluding LUCF" = "#33A02C",
    "Total including LUCF" = "#FB9A99",
    "Transportation" = "#1B9E77"
  )) +
  labs(title = "Sectoral Contributions to Emissions (2021)",
       x = "Country",
       y = "Emissions (MtCO₂e)") +
  theme(legend.position = "right")
```
We can infer that for most of the countries in 2021, sectors including Energy and Electricity and heat are the top sectors that emit Co2.
Another, important observation to note is the bars below 0, that is Land Use Change and Forestry denotes that this sector acts as sink of Co2 rather than
source and thus, the negative value.


4. Next we examine the trends of emissions by specific regions. 
```{r}
facet_data <- subset(data_long, Country %in% c("China", "United States", "India", "European Union (27)","Russia") & Sector == "Total including LUCF")

ggplot(facet_data, aes(x = Year, y = Emissions, color = Gas)) +
  geom_line(size = 1) +
  facet_wrap(~ Country, scales = "free_y") +
  labs(title = "Regional Emissions Trends by Gas",
       x = "Year",
       y = "Emissions (MtCO₂e)") +
  theme_minimal()
```
We observe that China , United States and India are the largest emitters of Co2 in recent years. Historically, United States , China and European Union have
been the highest emitters of CO2. Moreover, EU and United States have shown some decline in emissions which is a good indication.  F-Gas and N2O are relatively small
for all the regions.

```{r}
facet_data <- subset(
  data_long,
  Country %in% c(
    "China", "United States", "India", "European Union (27)",
    "Russia", "Japan", "Brazil", "Indonesia", "Iran", "Germany" 
  ) &
    Sector %in% c(
      "Energy", "Agriculture", "Transport",
      "Electricity/Heat", "Waste", "Industrial Processes"
    ) &  
    Gas %in% c("CO2", "CH4", "N2O", "F-Gas")  
)
```

5. For this plot we analyze the trend in the emissions of the gases in the main sectors over the years.
```{r}
gas_sector_data <- subset(
  data_long,
  Country == "World" & Sector %in% c("Energy", "Agriculture", "Electricity/Heat") & Gas %in% c("CO2", "CH4", "N2O", "F-Gas")
)

ggplot(gas_sector_data, aes(x = Year, y = Emissions, color = Gas)) +
  geom_line(size = 1) +
  facet_wrap(~ Sector, scales = "free_y") +
  labs(
    title = "Gas Contributions by Sector Over Time (1990–2021)",
    x = "Year",
    y = "Emissions (MtCO₂e)",
    color = "Gas"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 15, face = "bold"),
    axis.text = element_text(size = 9),
    legend.position = "right"
  )
```
An interesting information that we get from this plot is that the agriculture sector is mainly responsible for emitting CH4 and and N2O. Whereas, the major 
contributors of CO2 have been Electricity/Heat and Energy Sectors. Moreover, over the years, the emissions of these gases have increased. Understanding these 
trends in these sectors can help policy makers develop effective measures to control emissions and adopt sustainable practices.

6. In this plot we analyze the trends of emission of different gases in different countries in different sectors.
```{r,fig.width=15,fig.height=10}

gas_focused <- subset(
  data_long,
  Country %in% c("China", "United States", "India", "European Union (27)", "Russia") &
    Sector %in% c("Energy", "Agriculture", "Transport", "Electricity/Heat") &
    Gas %in% c("CO2", "CH4", "N2O")
)


ggplot(gas_focused, aes(x = Year, y = Emissions, color = Country)) +
  geom_line(size = 1) +
  facet_grid(Sector ~ Gas, scales = "free_y") +
  labs(
    title = "Sectoral Emissions by Gas and Country (1990–2021)",
    x = "Year",
    y = "Emissions (MtCO₂e)",
    color = "Country"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10, face = "bold"),
    plot.title = element_text(size = 16, face = "bold"),
    legend.position = "bottom"
  )

```
China and United States are the primary emitters of CO2 mainly in the Energy and Electricity/ Heat Sectors. Further, The European Union suggests a stable
trend in the emission of CO2, with sector agriculture showing a downward trend in the emission of CH4 and N2O. Russia also displays a similar trend India 
leads the emission of CH4 particularly in the agriculture sector. This can be attributed to the large number of agricultural activities carried out in India.
United States also displays a stable trend in agriculture sector and slight downward trend in the Electricity and Energy sector.

7. Sectoral Emissions across countries in 2021 or CO2
```{r,fig.width=10,fig.height=5}
sector_country_data <- subset(
  data_long,
  Year == 2021 &
    Country %in% c(
      "China", "United States", "India", "European Union (27)",
      "Russia", "Japan", "Brazil", "Germany"
    ) &
    Gas == "CO2" &
    !Sector %in% c("Total including LUCF", "Total excluding LUCF")  # Exclude these sectors
)

ggplot(sector_country_data, aes(x = Country, y = Sector, fill = Emissions)) +
  geom_tile() +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  labs(
    title = "Sectoral Emissions Across Countries (2021)",
    x = "Country",
    y = "Sector",
    fill = "Emissions (MtCO₂e)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right"
  )
```
This plot informs us about the emissions of each country with respect to a particular sector. This also tells us how active that country would be in that
sector. As we observed from the previous graphs, Energy and Electricity/Heat are the top emitters. Couple of countries have high amount of emission in the
transportation and manufacturing/construction. 

8. Here we analyze the sectoral emissions by gas which a plot between two categorical varibles and thus we use a stacked bar chart.
```{r}
# Filter 
sector_gas_data <- subset(
  data_long,
  Year == 2021 &
    !Sector %in% c("Total including LUCF", "Total excluding LUCF") &
    Gas %in% c("CO2", "CH4", "N2O", "F-Gas")
)
```

```{r}

sector_gas_summary <- sector_gas_data |>
  group_by(Sector, Gas) |>
  summarise(Total_Emissions = sum(Emissions, na.rm = TRUE)) |>
  ungroup()


ggplot(sector_gas_summary, aes(x = reorder(Sector, -Total_Emissions), y = Total_Emissions, fill = Gas)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Sectoral Emissions by Gas (2021)",
    x = "Sector",
    y = "Total Emissions (MtCO₂e)",
    fill = "Gas"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right"
  )
```
Here we observe that Energy, Electricity/Heat, Transportation, Manufacturing/Construction are the major contributors of CO2 emissions. Sectors like Agriculture,
Fugitive emissions and waste contribute to the emissions of CH4. Further Agriculture and Industrial sectors are responsible for the emissions of N2O gas. And Industrial
processes are the major contributors of F-Gas.

9. In this plot we finally analyze the percentage contribution of the top 5 CO2 emitting countries over the years.
```{r}
library(dplyr)
library(ggplot2)

# Filter CO2 emissions data
co2_data <- data_long |>
  filter(Gas == "CO2", Sector == "Total including LUCF")

# Calculate global total emissions for each year
global_emissions <- co2_data |>
  group_by(Year) |>
  summarise(Total_Emissions = sum(Emissions, na.rm = TRUE))

# Calculate cumulative emissions per country and identify top 5 emitters
top_5_countries <- co2_data |>
  group_by(Country) |>
  summarise(Cumulative_Emissions = sum(Emissions, na.rm = TRUE)) |>
  arrange(desc(Cumulative_Emissions)) |>
  slice(2:6) |>
  pull(Country)

# Filter for top 5 countries and calculate percentage contribution
top_5_contributions <- co2_data |>
  filter(Country %in% top_5_countries) |>
  left_join(global_emissions, by = "Year") |>
  mutate(Percentage_Contribution = Emissions / Total_Emissions * 100)

# Plot the line chart
ggplot(top_5_contributions, aes(x = Year, y = Percentage_Contribution, color = Country)) +
  geom_line(size = 1) +
  labs(
    title = "Contribution of Top 5 Countries to CO2 Emissions (1990–2021)",
    x = "Year",
    y = "Percentage Contribution",
    color = "Country"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 10),
    legend.position = "right"
  )

```
We observe that the percentage contribution of China has increased significantly over the years. The same trend is observed for India but the rise is gradual and is 
not as steep as China. Whereas, for the countries in European Union, United States and Brazil, the contribution of CO2 has fallen over the years.

```{r}
library(dplyr)
library(ggplot2)

# Filter CO2 emissions data
co2_data <- data_long |>
  filter(Gas == "CO2", Sector == "Total including LUCF")

# Calculate global total emissions for each year
global_emissions <- co2_data |>
  group_by(Year) |>
  summarise(Total_Emissions = sum(Emissions, na.rm = TRUE))

# Calculate cumulative emissions per country and identify top 5 emitters
top_5_countries <- co2_data |>
  group_by(Country) |>
  summarise(Cumulative_Emissions = sum(Emissions, na.rm = TRUE)) |>
  arrange(desc(Cumulative_Emissions)) |>
  slice(2:6) |>
  pull(Country)

# Separate emissions into top 5 countries and other countries
stacked_data <- co2_data |>
  left_join(global_emissions, by = "Year") |>
  mutate(
    Group = ifelse(Country %in% top_5_countries, "Top 5 Countries", "Other Countries"),
    Percentage_Contribution = Emissions / Total_Emissions * 100
  ) |>
  group_by(Year, Group) |>
  summarise(Total_Percentage = sum(Percentage_Contribution, na.rm = TRUE)) |>
  ungroup()

# Plot the stacked bar chart
ggplot(stacked_data, aes(x = Year, y = Total_Percentage, fill = Group)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(
    title = "Contribution of Top 5 Countries vs Other Countries to CO2 Emissions (1990–2021)",
    x = "Year",
    y = "Percentage Contribution",
    fill = "Group"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 10),
    legend.position = "right"
  )

```
From 1990 to 2021, we can see the contribution of top 5 countries to CO2 Emissions increases less than 25% to slightly more than 25%. Noticeably, there is a large decrease from 2010 to 2011. After that, the percentage contribution from the top 5 countries stabilizes. 

