# Results

```{r}
library(readr)
library(tidyverse)
```

```{r}
data <- read.csv("Datasets/historical_emissions.csv", na.strings = c("NA", "N/A"))

tail(data)
dim(data)
```

```{r}
cleaned_data <- na.omit(data)
dim(cleaned_data)
str(cleaned_data)
data_long <- cleaned_data |>
  pivot_longer(
    cols = starts_with("X"),    # Select all columns starting with "X" (years)
    names_to = "Year",          # New column for years
    values_to = "Emissions"     # New column for emission values
  ) |>
  mutate(Year = as.numeric(sub("X", "", Year)))
data_long
```
#should we drop data source column?

```{r}
# Filter for global emissions across all gases for "Total including LUCF"
global_emissions_gases <- subset(data_long, Country == "World" & Sector == "Total including LUCF")

# Plot emissions for all gases with different line colors
ggplot(global_emissions_gases, aes(x = Year, y = Emissions, color = Gas)) +
  geom_line(size = 1) +
  scale_y_continuous(trans = 'log10', labels = scales::comma) +
  labs(title = "Global Emissions Over Time by Gas",
       x = "Year",
       y = "Emissions (MtCO₂e)(on log scale)",
       color = "Gas") +
  theme_minimal() +
  theme(legend.position = "bottom")

global_emissions_gases <- subset(data_long, Country == "World" & Sector == "Total including LUCF")
```

```{r}
# Plot emissions for all gases with different line colors
ggplot(global_emissions_gases, aes(x = Year, y = Emissions, color = Gas)) +
  geom_line(size = 1) +
  labs(title = "Global Emissions Over Time by Gas",
       x = "Year",
       y = "Emissions (MtCO₂e)(on log scale)",
       color = "Gas") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
data_2021 <- subset(data_long, Year == 2021 & Gas == "CO2" & Sector == "Total including LUCF")
top_10 <- data_2021[order(-data_2021$Emissions), ][2:11, ]

ggplot(top_10, aes(x = reorder(Country, Emissions), y = Emissions)) +
  geom_bar(stat = "identity", fill = "cornflowerblue") +
  scale_y_continuous(breaks = seq(0, max(top_10$Emissions), by = 1000), labels = scales::comma) +
  labs(title = "Top 10 Emitting Countries in 2021",
       x = "Country",
       y = "Emissions (MtCO₂e)") +
  theme_minimal() +
  coord_flip()
```

```{r}
top_countries <- top_10$Country
sector_data <- subset(data_long, Year == 2021 & Country %in% top_countries & Gas == "CO2")

ggplot(sector_data, aes(x = Country, y = Emissions, fill = Sector)) +
  geom_bar(stat = "identity") +
  labs(title = "Sectoral Contributions to Emissions (2021)",
       x = "Country",
       y = "Emissions (MtCO₂e)") +
  theme_minimal() +
  theme(legend.position = "bottom")
```
## Changes made: change Total including LUCF Total excluding LUCF to LUCF and Total excluding LUCF, wondering if that makes sense. Should we apply this to the whole dataset? Also, should we change the color scale, to make each sector more distinct, as some colors stay together has small difference. Besides, there seem to be negative values(below 0), wondering why this is happenning.
```{r}
# Calculate LUCF emissions
sector_data <- subset(data_long, Year == 2021 & Country %in% top_countries & Gas == "CO2")

# Create a new LUCF column by calculating the difference
sector_data_lucf <- sector_data |>
  filter(Sector %in% c("Total including LUCF", "Total excluding LUCF")) |>
  spread(Sector, Emissions) |>
  mutate(LUCF = `Total including LUCF` - `Total excluding LUCF`) |>
  gather(key = "Sector", value = "Emissions", c("LUCF", "Total excluding LUCF"))

# Combine with other sectors
sector_data_combined <- sector_data |>
  filter(!Sector %in% c("Total including LUCF", "Total excluding LUCF")) |>
  bind_rows(sector_data_lucf)

# Plot the updated data
ggplot(sector_data_combined, aes(x = Country, y = Emissions, fill = Sector)) +
  geom_bar(stat = "identity") +
  labs(title = "Sectoral Contributions to Emissions (2021)",
       x = "Country",
       y = "Emissions (MtCO₂e)") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_brewer(palette = "PuBu") 
```



```{r}
facet_data <- subset(data_long, Country %in% c("China", "United States", "India", "European Union (27)") & Sector == "Total including LUCF")

ggplot(facet_data, aes(x = Year, y = Emissions, color = Gas)) +
  geom_line(size = 1) +
  facet_wrap(~ Country, scales = "free_y") +
  labs(title = "Regional Emissions Trends by Gas",
       x = "Year",
       y = "Emissions (MtCO₂e)") +
  theme_minimal()
```

```{r}
facet_data <- subset(
  data_long,
  Country %in% c(
    "China", "United States", "India", "European Union (27)",
    "Russia", "Japan", "Brazil", "Indonesia", "Iran", "Germany"  # Top 10 emitters
  ) &
    Sector %in% c(
      "Energy", "Agriculture", "Transport",
      "Electricity/Heat", "Waste", "Industrial Processes"
    ) &  # Key sectors
    Gas %in% c("CO2", "CH4", "N2O", "F-Gas")  # Major greenhouse gases
)
```


```{r}
gas_sector_data <- subset(
  data_long,
  Country == "World" & Sector %in% c("Energy", "Agriculture", "Transport", "Electricity/Heat") & Gas %in% c("CO2", "CH4", "N2O", "F-Gas")
)

ggplot(gas_sector_data, aes(x = Year, y = Emissions, color = Gas)) +
  geom_line(size = 1) +
  facet_wrap(~ Sector, scales = "free_y") +
  labs(
    title = "Gas Contributions by Sector Over Time (1990–2021)",
    x = "Year",
    y = "Emissions (MtCO₂e)",
    color = "Gas"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10, face = "bold"),
    plot.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 8),
    legend.position = "bottom"
  )
```

# Should we free scale in this part?
```{r}
# Filter data for selected countries and sectors across gases
gas_focused <- subset(
  data_long,
  Country %in% c("China", "United States", "India", "European Union (27)", "Russia") &
    Sector %in% c("Energy", "Agriculture", "Transport", "Electricity/Heat") &
    Gas %in% c("CO2", "CH4", "N2O", "F-Gas")
)

# Faceted grid plot
ggplot(gas_focused, aes(x = Year, y = Emissions, color = Country)) +
  geom_line(size = 1) +
  facet_grid(Sector ~ Gas, scales = "free_y") +
  labs(
    title = "Sectoral Emissions by Gas and Country (1990–2021)",
    x = "Year",
    y = "Emissions (MtCO₂e)",
    color = "Country"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 10, face = "bold"),
    plot.title = element_text(size = 16, face = "bold"),
    legend.position = "bottom"
  )

```


```{r}
sector_country_data <- subset(
  data_long,
  Year == 2021 &
    Country %in% c(
      "China", "United States", "India", "European Union (27)",
      "Russia", "Japan", "Brazil", "Germany"
    ) &
    Gas == "CO2" &
    !Sector %in% c("Total including LUCF", "Total excluding LUCF")  # Exclude these sectors
)

ggplot(sector_country_data, aes(x = Country, y = Sector, fill = Emissions)) +
  geom_tile() +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  labs(
    title = "Sectoral Emissions Intensity Across Countries (2021)",
    x = "Country",
    y = "Sector",
    fill = "Emissions (MtCO₂e)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right"
  )
```



```{r}
# Filter data for sector-by-gas analysis
sector_gas_data <- subset(
  data_long,
  Year == 2021 &
    !Sector %in% c("Total including LUCF", "Total excluding LUCF") &
    Gas %in% c("CO2", "CH4", "N2O", "F-Gas")
)
```

```{r}
# Summarize data to get total emissions by sector and gas
sector_gas_summary <- sector_gas_data |>
  group_by(Sector, Gas) |>
  summarise(Total_Emissions = sum(Emissions, na.rm = TRUE)) |>
  ungroup()

# Create stacked bar chart
ggplot(sector_gas_summary, aes(x = reorder(Sector, -Total_Emissions), y = Total_Emissions, fill = Gas)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Sectoral Emissions by Gas (2021)",
    x = "Sector",
    y = "Total Emissions (MtCO₂e)",
    fill = "Gas"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
```


```{r}
library(ggplot2)

# Filter data for relevant countries, sectors, and gases
grouped_bar_data <- subset(
  data_long,
  Year == 2021 &
    Country %in% c("China", "United States", "India", "European Union (27)", "Russia") &
    !Sector %in% c("Total including LUCF", "Total excluding LUCF") &
    Gas %in% c("CO2", "CH4", "N2O","F-Gas")
)

# Create grouped bar chart
ggplot(grouped_bar_data, aes(x = Sector, y = Emissions, fill = Gas)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ Country, scales = "free_y") +
  labs(
    title = "Sectoral Emissions by Gas Faceted by Country (2021)",
    x = "Sector",
    y = "Emissions (MtCO₂e)",
    fill = "Gas"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
```

```{r}
library(ggplot2)

# Filter data for relevant countries, sectors, and gases
grouped_bar_data <- subset(
  data_long,
  Year == 2021 &
    Country %in% c("China", "United States", "India", "European Union (27)", "Russia") &
    !Sector %in% c("Total including LUCF", "Total excluding LUCF") &
    Gas %in% c("CO2", "CH4", "N2O")
)

# Create grouped bar chart
ggplot(grouped_bar_data, aes(x = Sector, y = Emissions, fill = Gas)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(trans = "log10", labels = scales::comma) +
  facet_wrap(~ Country, scales = "free_y") +
  labs(
    title = "Sectoral Emissions by Gas Faceted by Country (2021)",
    x = "Sector",
    y = "Emissions (MtCO₂e)",
    fill = "Gas"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
```

# Additional Graph
```{r}
library(dplyr)
library(ggplot2)

# Filter CO2 emissions data
co2_data <- data_long |>
  filter(Gas == "CO2", Sector == "Total including LUCF")

# Calculate global total emissions for each year
global_emissions <- co2_data |>
  group_by(Year) |>
  summarise(Total_Emissions = sum(Emissions, na.rm = TRUE))

# Calculate cumulative emissions per country and identify top 5 emitters
top_5_countries <- co2_data |>
  group_by(Country) |>
  summarise(Cumulative_Emissions = sum(Emissions, na.rm = TRUE)) |>
  arrange(desc(Cumulative_Emissions)) |>
  slice(2:6) |>
  pull(Country)

# Filter for top 5 countries and calculate percentage contribution
top_5_contributions <- co2_data |>
  filter(Country %in% top_5_countries) |>
  left_join(global_emissions, by = "Year") |>
  mutate(Percentage_Contribution = Emissions / Total_Emissions * 100)

# Plot the line chart
ggplot(top_5_contributions, aes(x = Year, y = Percentage_Contribution, color = Country)) +
  geom_line(size = 1) +
  labs(
    title = "Contribution of Top 5 Countries to CO2 Emissions (1990–2021)",
    x = "Year",
    y = "Percentage Contribution",
    color = "Country"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 10),
    legend.position = "right"
  )

```



